Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-08-21 Master Thesis/synopsis.tex
Fix second batch typo

For the second assertion, let $K$ be not algebraically closed and $v=v(w_1)=v(w_2)\in\operatorname{Spv}_\text{d}^{(1)}(K)$. The restriction $v|$ of $v$ on $k=\mathbb{Q}(T)\subset K$ is a discrete valuation. Choose a henselization $k_{v|}$ of $k$ such that $k_{v|}\subset K$ and consider the absolute inertia field $k_{v|}^\text{t}$ of $k_{v|}$. Since $k(v)$ is separably closed, we have $k_{v|}^\text{t}\subset K$, i.e. $G_K\subset G_{k_{v|}^\text{t}}$. Since $v|\in \operatorname{Spv}_\text{d}^{(1)} ({k_{v|}^\text{t}})$ is discrete and henselian, \cite{Ka79} §5, Theorem 1 shows $\operatorname{cd}\big(G_{k_{v|}^\text{t}}\big)\leq 2$. In particular, $\operatorname{cd}(G_K)\leq 2$.
\end{proof}

@@

\begin{proof}
Just like in Lemma \ref{2.4-finitelygenerated} we can assume that $\mu_\ell\subset K$ and $K$ is not formally real and replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2)$. We adopt the notation from the previous proofs.

@@

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. Then the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
The implication (i)$\implies$(ii) follows from Lemma \ref{2.2}. Now assume (ii) holds. Let $v=v(w_1)\in \operatorname{Spv}_\text{d}^{(1)}(K_{w_1}K_{w_2})$. Then $k(v)$ is separably closed. We have $K_{w_1}K_{w_2}\subset K_v^\text{t}$ and hence $K_{w_1}K_{w_2}=K_v^\text{t}$ by Theorem \ref{theorem-absoluteinertia}. $\Gamma_v$ is equal to the value group of $v|_K\in\operatorname{Spv}_\text{d}^{(1)}(K)$, which is finitely generated. However, value groups of valuations on an algebraically closed field are divisible and in particular not finitely generated, hence $K_v^\text{t}$ can't be algebraically closed and neither can the composite $K_{w_1}K_{w_2}$.

@@

Let $\mathcal{P}(F)$ denote the group of principal ideals of $F$, $\mathcal{I}(F)$ the ideal group of $F$, and $\operatorname{Cl}(F) = \mathcal{I}(F)/\mathcal{P}(F)$ the ideal class group of $F$. We have the following exact sequence:
\[ \begin{tikzcd}
1 \ar[r]& \mathcal{O}_{F_1}^\times /\mathcal{O}_F^\times\ar[r] & F_1^\times / F^\times\ar[r] & \mathcal{P}(F_1)/\mathcal{P}(F)\ar[r] & 1
\end{tikzcd} \]
Again by Lemma \ref{5.2} (ii), it suffices to show that $\mathcal{P}(F_1)/\mathcal{P}(F)$ is free-by-finite. 
Consider the following diagram with exact rows:
\[ \begin{tikzcd}
0\ar[r] & \mathcal{P}(F) \ar[d] \ar[r]& \mathcal{I}(F) \ar[r]\ar[d]& \operatorname{Cl}(F)\ar[r]\ar[d, "f"] & 0\\
0 \ar[r]& \mathcal{P}(F_1) \ar[r]& \mathcal{I}(F_1) \ar[r]& \operatorname{Cl}(F_1) \ar[r]& 0 
\end{tikzcd} \]
Using the snake lemma, we get the following exact sequence:
\[
0 \longrightarrow \ker(f)\longrightarrow \mathcal{P}(F_1)/\mathcal{P}(F)\longrightarrow \mathcal{I}(F_1)/\mathcal{I}(F)
\longrightarrow \operatorname{coker}(f)\longrightarrow 0 \]
Since the class numbers of number fields are finite, according to Lemma \ref{5.2} (i), it suffices to show that $\mathcal{I}(F_1)/\mathcal{I}(F)$ is free-by-finite. This follows from the fact that there are only finitely many primes of $F$ which are ramified in $F_1$.\qedhere
\end{enumerate}
\end{proof}

@@

\begin{corollary}\label{mordell-weil-corollary}
Let $C$ be a smooth projective curve over a number field $F$. Then the Picard group $\operatorname{Pic}(C)$ is finitely generated.
\end{corollary}

@@

\begin{proof}
Observe that $\varphi(-1)^2 = \varphi(1) = 1$, and since $\varphi$ is bijective on $K'^\times$, it follows $\varphi(-1) \neq 1$ and hence $\varphi(-1) = -1$. Now, to prove the assertion, it suffices to show $\varphi(x + y) = \varphi(x) + \varphi(y)$ for all $x,y\in K'^\times$.
\begin{itemize}
\item \textit{Case 1:} If $x + y = 0$, we have:
\[ \varphi(x + y) = 0 = \varphi(x) + \varphi(-1)\varphi(x) = \varphi(x) + \varphi(-x) = \varphi(x) + \varphi(y) \]
\item \textit{Case 2:} Let $x + y\neq 0$ and $S\subset X_0$ be an admissible finite subset such that $\varphi(x + y)$, $\varphi(x)$ and $\varphi(y)$ are contained in $K^S$. Looking at the commutative diagram in \mbox{Step 1} we get for all $P\in X_0 -  S$:
\begin{align*}
r_P(\varphi(x+ y)) &= \varphi_P(r_{P'}(x + y)) \\
&= \varphi_P(r_{P'}(x)) + \varphi_P(r_{P'}(y)) = r_P(\varphi(x) + \varphi(y))
\end{align*}
Note that $\varphi_P: k(P')\to k(P)$ is a field isomorphism and that the residue maps $r_{P'}: \mathcal{O}_{P'} \to k(P')$ are ring homomorphisms. Since $K^S\to \prod_{P\in X_0 -  S} k(P)^\times$ is injective, we get $\varphi(x+y) = \varphi(x) + \varphi(y)$.\qedhere
\end{itemize}
\end{proof}

>>>>

For the second assertion, let $K$ be not algebraically closed and $v=v(w_1)=v(w_2)\in\operatorname{Spv}_\text{d}^{(1)}(K)$. The restriction $v|$ of $v$ on $k=\mathbb{Q}(T)\subset K$ is a discrete valuation since $k$ is finitely generated. Choose a henselization $k_{v|}$ of $k$ such that $k_{v|}\subset K$ and consider the absolute inertia field $k_{v|}^\text{t}$ of $k_{v|}$. Since $k(v)$ is separably closed, we have $k_{v|}^\text{t}\subset K$, i.e. $G_K\subset G_{k_{v|}^\text{t}}$. Since $v|\in \operatorname{Spv}_\text{d}^{(1)} ({k_{v|}^\text{t}})$ is discrete and henselian, \cite{Ka79} §5, Theorem 1 shows $\operatorname{cd}\big(G_{k_{v|}^\text{t}}\big)\leq 2$. In particular, $\operatorname{cd}(G_K)\leq 2$.
\end{proof}

@@

\begin{proof}
Just like in Lemma \ref{2.4-finitelygenerated} we can assume that $\mu_\ell\subset K$ and $K$ is not formally real. Furthermore, we replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2)$. We adopt the notation from the previous proofs.

@@

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. Then the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
The implication (i)$\implies$(ii) follows from Lemma \ref{2.2}. Now assume (ii) holds and let $v=v(w_1)\in\operatorname{Spv}^{(1)}_\text{d}(K)$. We have $K_{w_1}K_{w_2}\subset K_v^\text{t}$. By Theorem \ref{theorem-absoluteinertia}, the value group of $v$ on $K_v^\text{t}$ is equal to the value group of $v$ on $K$, which is finitely generated by Lemma \ref{rank-dimension-inequality}. However, value groups of valuations on an algebraically closed field are divisible and in particular not finitely generated, hence $K_v^\text{t}$ can't be algebraically closed and neither can the composite $K_{w_1}K_{w_2}$.

Furthermore, if (ii) holds, the extension of $v$ on $K_{w_1}K_{w_2}$ has a separably closed residue field. By Theorem \ref{theorem-absoluteinertia}, we have $K_{w_1}K_{w_2}=K_v^\text{t}$.

@@

Let $\mathcal{P}(F)$ denote the group of principal ideals of $F$, $\mathcal{I}(F)$ the ideal group of $F$, and $\operatorname{Cl}(F) = \mathcal{I}(F)/\mathcal{P}(F)$ the ideal class group of $F$. We have the following exact sequence:
\[ \begin{tikzcd}
1 \ar[r]& \mathcal{O}_{F_1}^\times /\mathcal{O}_F^\times\ar[r] & F_1^\times / F^\times\ar[r] & \mathcal{P}(F_1)/\mathcal{P}(F)\ar[r] & 1
\end{tikzcd} \]
Again by Lemma \ref{5.2} (ii), it suffices to show that $\mathcal{P}(F_1)/\mathcal{P}(F)$ is free-by-finite. 
Consider the following commutative diagram with exact rows:
\[ \begin{tikzcd}
0\ar[r] & \mathcal{P}(F) \ar[d] \ar[r]& \mathcal{I}(F) \ar[r]\ar[d]& \operatorname{Cl}(F)\ar[r]\ar[d, "f"] & 0\\
0 \ar[r]& \mathcal{P}(F_1) \ar[r]& \mathcal{I}(F_1) \ar[r]& \operatorname{Cl}(F_1) \ar[r]& 0 
\end{tikzcd} \]
Using the snake lemma, we get the following exact sequence:
\[
0 \longrightarrow \ker(f)\longrightarrow \mathcal{P}(F_1)/\mathcal{P}(F)\longrightarrow \mathcal{I}(F_1)/\mathcal{I}(F)
\longrightarrow \operatorname{coker}(f)\longrightarrow 0 \]
Since the class numbers of number fields are finite, according to Lemma \ref{5.2} (i), it suffices to show that $\mathcal{I}(F_1)/\mathcal{I}(F)$ is free-by-finite. This follows from the fact that there are only finitely many primes of $F$ which are ramified in $F_1$.\qedhere
\end{enumerate}
\end{proof}

@@

\begin{corollary}\label{mordell-weil-corollary}
Let $C$ be a smooth projective curve over a number field $F$. Then its Picard group $\operatorname{Pic}(C)$ is finitely generated.
\end{corollary}

@@

\begin{proof}
Observe that $\varphi(-1)^2 = \varphi(1) = 1$, and since $\varphi$ is bijective on $K'^\times$, it follows $\varphi(-1) \neq 1$ and hence $\varphi(-1) = -1$. Now, to prove the assertion, it suffices to show $\varphi(x + y) = \varphi(x) + \varphi(y)$ for all $x,y\in K'^\times$.
\begin{itemize}
\item \textit{Case 1:} If $x + y = 0$, we have:
\[ \varphi(x + y) = 0 = \varphi(x) + \varphi(-1)\varphi(x) = \varphi(x) + \varphi(-x) = \varphi(x) + \varphi(y) \]
\item \textit{Case 2:} Let $x + y\neq 0$ and $S\subset X_0$ be an admissible finite subset such that $\varphi(x + y)$, $\varphi(x)$ and $\varphi(y)$ are contained in $K^S$. Looking at the commutative diagram in \mbox{Step 1} we get for all $P\in X_0 -  S$:
\begin{align*}
r_P(\varphi(x+ y)) &= \varphi_P(r_{P'}(x + y)) \\
&= \varphi_P(r_{P'}(x)) + \varphi_P(r_{P'}(y)) = r_P(\varphi(x) + \varphi(y))
\end{align*}
Note that $\varphi_P: k(P')\to k(P)$ is a field isomorphism and that the residue maps $r_{P'}: \mathcal{O}_{P'} \to k(P')$ are ring homomorphisms. Since $K^S\to \prod_{P\in X_0 -  S} k(P)^\times$ is injective by Lemma \ref{5.4}, we get $\varphi(x+y) = \varphi(x) + \varphi(y)$.\qedhere
\end{itemize}
\end{proof}
